---
title: "DSC week 14 IP - Association Rules and Anomaly Detection"
author: "Wambui"
date: "29/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



##  CarreFour Kenya Sales

### 1. Research Question
Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest number of sales (total price including tax).


### 2. Metric of Success

identifying the Association Rules and Anomaly Detection.

### 3. Understanding the Context.

Sales transactional data has been acquired from the Carrefour Kenya stores in Kenya 

### 4. Recording the Experimental Design

a. Data Loading
b. Data Cleaning and processing
c. Exploratory Data Analysis
d. Association Rules and Anomaly Detection
e. Recommendations and Conclusions

### 5. Data Relevance.

The provided data is relevant for this kind of study 

### Data Preview


```{r}
install.packages("arules")
install.packages("data.table")
```


```{r}
#
#loading the library
library(data.table) # load package
library(tidyverse)
library("caret")
library(arules)
```


```{r}

# Dataset Url = http://bit.ly/SupermarketDatasetII
# Importing our dataset
# 
#
sales <- read.csv('http://bit.ly/SupermarketDatasetII')
head(sales)
```



# Check the Data
```{r}
head(sales,6)

```

```{r}
tail(sales,6)
```


Converting entries to transactions.
```{r}
library(arules)
path <-"http://bit.ly/SupermarketDatasetII"
Transactions<-read.transactions(path, sep = ",")
Transactions
```

Verifying the object's class
```{r}
class(Transactions)
```


```{r}

# Previewing our first 5 transactions
#
inspect(Transactions[1:5])
```
### Previewing the dataset
```{r}
items<-as.data.frame(itemLabels(Transactions))
colnames(items) <- "Item"
head(items, 10) 
```
Observation: There are 10 items that make up our dataset

### checking the dataset summary

```{r}
summary(Transactions)

```
Observation: There are 8 numerical columns and 8 categorical columns.


### Checking the Frequency of some articles 
# of transactions ranging from 1 to 5
```{r}
itemFrequency(Transactions[, 1:5],type = "absolute")
round(itemFrequency(Transactions[, 1:5],type = "relative")*100,2)

```

# Displaying top 10 most common items in the transactions dataset 
# and the items whose relative importance is at least 10%
# 
```{r}
par(mfrow = c(1, 2))

# plot the frequency of items
itemFrequencyPlot(Transactions, topN = 10,col="darkgreen")
itemFrequencyPlot(Transactions, support = 0.1,col="darkred")
```

### Building the Model

#model1 : support =0.001 and confidence = 0.8
```{r}
sales.rules <- apriori (Transactions, parameter = list(supp = 0.001, conf = 0.8))
sales.rules
```

We build Model1 using association rules and apriori() function, min support=0.001, confidence=0.8
Observation: Model 1 has 74 rules.

### increasing the support and lowering the confidence level
```{r}
# Building a apriori model with Min Support as 0.002 and confidence as 0.8.
sales.rules2<- apriori (Transactions,parameter = list(supp = 0.002, conf = 0.8)) 
```



```{r}
# Building a apriori model with Min Support as 0.002 and confidence as 0.6.
sales.rules2<- apriori (Transactions,parameter = list(supp = 0.002, conf = 0.6)) 
```


Exploring our model
```{r}
summary(sales.rules)
```



##Observing rules built in our model
```{r}
inspect(sales.rules[1:5])
```
Observation:

 If a customer buys frozen smoothie and spinach, they are 88% likely to buy mineral water also
 
 If a customer  buys bacon and pancake, they are 81% likely to buy spaghetti also.
 
 If a customer  buys mushroom cream sauce, and pasta, they are 95% likely to buy escalope also
  


```{r}
library(arulesViz)
plot(sales.rules,method = "graph",)
```


### Ordering these rules by criteria
```{r}
# Ordering these rules by a criteria such as the level of confidence
# then looking at the first five rules.
# We can also use different criteria such as: (by = "lift" or by = "support")
# 
sales.rules<-sort(sales.rules, by="confidence", decreasing=TRUE)
inspect(sales.rules[1:5])
```


The above 4 rules have a confidence of 100.

###. checking the top 5 items that will be bought before buying spagetti

```{r}
# What if we wanted to determine items that customers might buy 
# who have previously bought yogurt?
# ---
# 
# Subset the rules
spags <- subset(sales.rules, subset = lhs %pin% "spaghetti")

# Order by confidence
spags<-sort(spags, by="confidence", decreasing=TRUE)

```



```{r}
# inspect top 5
inspect(spags[1:5])
```



### Part 4: Anomaly Detection

# Dataset Url = http://bit.ly/SupermarketDatasetII
# Importing our dataset
# 
#
```{r}
sales_anom <- read.csv('http://bit.ly/CarreFourSalesDataset')
head(sales_anom)
```
# Check the Data
```{r}
head(sales_anom,6)

```


```{r}
tail(sales_anom,6)
```

```{r}
colnames(sales_anom)
```
Observation: The dataset contains 2 Columns namely Date and Sales
```{r}
dim(sales_anom)
```
Observation: The dataset contains 1000 rows and 2 Columns


### Convert date to date time data
```{r}
#changing date to date time. 
sales_anom$Date <- as.Date(sales_anom$Date, "%m/%d/%y")
head(sales_anom)
```

```{r}
# Install Lubridate package for date time data
#install.packages('lubridate')
library(lubridate)
library(tidyverse)
library(anomalize)
```

### Detecting Anomalies


```{r}
sales_tbl <- as_tibble(sales_anom)
head(sales_tbl)
```

```{r}
sales_anom_df <- sales_anom %>% group_by(Date) %>% tally()
colnames(sales_anom_df) <- c('TransactionDate', 'Total')
head(sales_anom_df)
```



```{r}
# we now plot using plot_anomaly_decomposition() to visualize out data.
sales_anom_df %>%
    time_decompose(Total) %>%
    anomalize(remainder) %>%
    plot_anomaly_decomposition(ncol = 2, alpha_dots = .8) +
    ggtitle("Anomaly Detection")
```
### plotting recomposition to check anomalies further
```{r}

sales_anom_df %>%
    time_decompose(Total) %>%
    anomalize(remainder) %>%
    time_recompose() %>%
    plot_anomalies(time_recomposed = TRUE, ncol = 4, alpha_dots = 0.25, fill="blue") +
    ggtitle("Anomaly Detection")
```
\


Observation: There are no anomaly seen.